---
output:
  github_document:
    html_preview: false
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# docker

The goal of docker is to provide access to the [docker](https://www.docker.com/) [API](https://docs.docker.com/engine/api/) via [reticulate](https://rstudio.github.io/reticulate/) using the [Python SDK](https://docker-py.readthedocs.io/en/stable/) for docker.

## Installation

```{r install, eval=FALSE}
if(!require(devtools)) {
  install.packages("devtools")
}
devtools::install_github('bhaskarvk/docker')
```

## Setup

Before you can use this package, you need to have python > 3.0 (preferably > 3.5), and the [docker](https://docker-py.readthedocs.io/en/stable/index.html) module which provides Python SDK for docker. A simple way to do this is using a [virtual environment](http://docs.python-guide.org/en/latest/dev/virtualenvs/). [`virtualenvwrapper`](http://docs.python-guide.org/en/latest/dev/virtualenvs/#virtualenvwrapper) makes setting up and working on virtual envs quite painless.

Once you have setup `virtualenvwrapper`, you need to create a new virtual environment and install the `docker` Python module in it.

```bash
mkvirtualenv --python=/usr/bin/python3 docker
workon docker
pip install docker

# Test the SDK againsts a locally running docker
# You should have a locally running docker for this.
python -c 'import docker; print(docker.from_env().version())'
```

You should see something like below, provided you had a locally running docker engine.

```
{'Os': 'linux', 'Arch': 'amd64', 'KernelVersion': '4.10.0-24-generic', 'GitCommit': '02c1d87', 'Version': '17.06.0-ce', 'BuildTime': '2017-06-23T21:19:04.990631145+00:00', 'MinAPIVersion': '1.12', 'GoVersion': 'go1.8.3', 'ApiVersion': '1.30'}
```

## Example


```{r hello-world}
reticulate::use_virtualenv("docker") # Here docker is the name of the python virtual env. we created above.
library(docker)
client <- docker$from_env()
client$containers$run("alpine", 'echo "Hello World!"')
```

## Usage

After you have a successful 'Hellow World!' displayed above, you can call every API supported by the Python SDK using reticulate. Please consult the documents for the SDK and the reticulate R package in the links given above.
