---
title: "Getting Started with 'docker'"
author: "user001 V. Karambelkar"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with 'docker'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The `docker` R package provides access to the [docker](https://www.docker.com/) [API](https://docs.docker.com/engine/api/) to programmatically control a docker engine from R. The docker engine could be running either locally or remotely. The `docker` package uses the [reticulate](https://rstudio.github.io/reticulate/) R package to invoke the [Python SDK](https://docker-py.readthedocs.io/en/stable/) for docker. Using the Python SDK allows the `docker` package to have a very minimal foot-print and yet allow complete docker API access. The package has been tested against docker engine running on [Linux](https://www.docker.com/docker-ubuntu) and [Microsoft Windows](https://www.docker.com/docker-windows) and should also work under [MacOS](https://www.docker.com/docker-mac).

Controlling a docker engine from withing R can be useful for ...

- Writing automated unit/integration tests for R packages against running services in docker containers. The package author can start a docker container, execute the tests, and tear the container down, all from withing the R code of the test.
- Another use case is easy access to containerized services as part of a researchers reproducible research workflow. Having access to a containerized service ensures that the research code executes identically regardless of where it's run.

## Installation

The package is not yet on CRAN so install it directly from GitHub using `devtools`.

```{r install, eval=FALSE}

install.packages('reticulate') # For Python access

if(!require(devtools)) {
  install.packages("devtools")
}
devtools::install_github('user001vk/docker')
```

## Setup

Before you can use this package you need to have Python > 3.0 (preferably > 3.5) with the [docker](https://docker-py.readthedocs.io/en/stable/index.html) Python module which provides Python SDK for docker. A simple way to do this is using a [virtual environment](http://docs.python-guide.org/en/latest/dev/virtualenvs/). [virtualenvwrapper](http://docs.python-guide.org/en/latest/dev/virtualenvs/#virtualenvwrapper) makes setting up Python modules quite painless.

Once you have installed Python 3.5+ and setup `virtualenvwrapper` you need to create a new virtual environment and install the `docker` Python module in it.

```bash
mkvirtualenv --python=/usr/bin/python3 docker
workon docker
pip install docker

# Test the SDK againsts a locally running docker
# You should have a locally running docker engine for this.
python -c 'import docker; print(docker.from_env().version())'
```

You should see something like below, provided you had a locally running docker engine.

```
{'Os': 'linux', 'Arch': 'amd64', 'KernelVersion': '4.10.0-24-generic', 'GitCommit': '02c1d87',
'Version': '17.06.0-ce', 'BuildTime': '2017-06-23T21:19:04.990631145+00:00',
'MinAPIVersion': '1.12', 'GoVersion': 'go1.8.3', 'ApiVersion': '1.30'}
```

## Examples

### Hello World!

Here we provide a simple getting started example. We start the [alpine](https://store.docker.com/images/alpine) docker image which is a small Linux image.We run a command to output "Hello World!" from within the started docker container and access it on the R side. 

```{r hello-world, eval=FALSE}
# This is a hint to the reticulate package to find a
# python environment with the "docker" python module available.
reticulate::use_virtualenv("docker")
library(docker)
client <- docker$from_env()
s <- client$containers$run("alpine", 'echo -n "Hello World!"', remove=TRUE)
print(s$decode("UTF-8"))
```

```
[1] Hello World!

```


### Real-World Example

Below is more of a real-world example where we start a [redis](https://redis.io/) container and interact with it using the `rredis` package, and subsequently tear it down.

```{r redis, eval=FALSE}
reticulate::py_discover_config("docker")
library(docker)
library(rredis)

docker_client <- docker$from_env()
docker_client$ping()

redis_instance <- docker_client$containers$run(
  'redis', name='redis', detach=TRUE, ports=list('6379/tcp'='6379'))
redis_instance$status

redisConnect(host='127.0.0.1', port=6379, nodelay=TRUE)
redisSet('docker-test', rnorm(20))
redisGet('docker-test')

redis_instance$stop()
redis_instance$remove()
```

And here's how it looks when run.

```
> reticulate::py_discover_config("docker")
python:         /home/user001/.virtualenvs/docker/bin/python
libpython:      /usr/lib/python3.5/config-3.5m-x86_64-linux-gnu/libpython3.5.so
pythonhome:     /usr:/usr
virtualenv:     /home/user001/.virtualenvs/docker/bin/activate_this.py
version:        3.5.2 (default, Nov 17 2016, 17:05:23)  [GCC 5.4.0 20160609]
numpy:          /home/user001/.virtualenvs/docker/lib/python3.5/site-packages/numpy
numpy_version:  1.13.1
docker:         /home/user001/.virtualenvs/docker/lib/python3.5/site-packages/docker

python versions found: 
 /home/user001/.virtualenvs/docker/bin/python
 /usr/bin/python
 /usr/bin/python3

> library(docker)

> library(rredis)

> docker_client <- docker$from_env()

> docker_client$ping()
[1] TRUE

> redis_instance <- docker_client$containers$run(
   'redis', name='redis', detach=TRUE, ports=list('6379/tcp'='6379'))

> redis_instance$status
[1] "created"

> Sys.sleep(5) # Wait for the container to start the service.

> redisConnect(host='127.0.0.1', port=6379, nodelay=TRUE)

> redisSet('docker-test', rnorm(20))
[1] "OK"

> redisGet('docker-test')
 [1]  1.01545119  0.08353996  0.71668645 -0.34256345 -0.09871241  0.24834106
 [7]  0.28802550  0.02321578 -1.62448884 -1.31099073  0.54530365  1.29799034
[13]  0.80973704  0.30018833 -0.72609402 -0.47678730 -0.79877246  1.08857741
[19] -1.92010832  1.61439356

> redis_instance$stop()
> redis_instance$remove()
```

The `reticulate::py_discover_config("docker")` call is used to ensure that `reticulate` was able to find a Python environment with the `docker` Python module installed.

### Other Docker APIs

You can call every API supported by the Python SDK using reticulate. Please consult the documents for the SDK and the reticulate R package in the links given above.

